---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# sansa

<!-- badges: start -->

<!-- badges: end -->

Machine learning is widely used in information-systems design. Yet, training algorithms on imbalanced datasets may severely affect performance on unseen data. For example, in some cases in healthcare, fintech, or cybersecurity contexts, certain subclasses are difficult to learn because they are underrepresented in training data. This R package offers a flexible and efficient solution based on a new synthetic average neighborhood sampling algorithm (SANSA), which, in contrast to other solutions, introduces a novel “placement” parameter that can be tuned to adapt to each dataset’s unique manifestation of the imbalance.

## Installation

You can install the released version of sansa from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("sansa")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("murtaza-nasir/sansa")
```

## Example

Lets first load some libraries.

```{r example}
library(sansa)
library(ggplot2)

```

Now lets generate an imbalanced dataset.

```{r dataset}

minority = data.frame(x1 = rnorm(50, 15, 2),
                      x2 = rnorm(50, 25, 10),
                      target = "true")
majority = data.frame(x1 = rnorm(500, 5, 4),
                      x2 = rnorm(500, 30, 10),
                      target = "false")

dataset = rbind(minority, majority)

ggplot(dataset) + geom_point(aes(x1, x2, color = target))
```

This imbalanced dataset can be balanced by SANSA using the `sansa` function.

```{r sansa}
sansaobject = sansa(x = dataset[,1:2], y = dataset$target, lambda = 1, ksel = 3)

balanced <- sansaobject$x
balanced$target = sansaobject$y

ggplot(balanced) + geom_point(aes(x1, x2, color = target))
```

SANSA returns a list object that can be used directly within the `caret` training pipeline.

## Details & Reference

Full details about the algorithm as well as benchmarks are available in the accompanying publication that will be added here shortly. 
